---
title: "Bioproject report"
params:
  prj: "PRJEB44932"
---

```{r}
#| include: false
#| label: setup
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
```

```{r}
#| label: load_libraries
suppressPackageStartupMessages({
    library(yaml)
    library(provoc)
    library(here)
    library(tidyverse)
})
all_data <- yaml.load_file(here("data", "data_sources.yml"))

bioproj <- all_data[[params$prj]]
```


# BioProject `r params$prj` - `r bioproj$author`


```{r}
#| label: load_data
filename <- here(
    "data/processed",
    paste0(params$prj, "_processed.csv.gz")
)

t0 <- Sys.time()
coco <- read.csv(filename)
coco$mutation <- parse_mutations(coco$label)
coco |>
    select(any_of(c("location", "date", "mutation", "count", "coverage"))) |>
    head() |>
    knitr::kable()
```


Loading data took `r round(difftime(Sys.time(), t0, units = "mins"), 2)` minutes.

## Summary Stats

From `data_sources.yml` and the loaded data:

- Claims `r bioproj$wwtps` locations, I see `r length(unique(coco$location))`.
- Claims `r bioproj$runs` runs, I see `r length(unique(coco$sra))`
- There are `r length(unique(coco$date))` unique dates.

```{r}
#| label: check_lineages
lineages_in_paper <- bioproj$lineages_in_paper
barcodes <- provoc::usher_barcodes()
not_in_barcodes <- setdiff(lineages_in_paper, rownames(barcodes))

available_lineages <- intersect(lineages_in_paper, rownames(barcodes))
barcodes <- filter_lineages(barcodes, available_lineages)

```

- Lineages in paper: `r lineages_in_paper`
- Lineages not in barcodes: `r not_in_barcodes`
- Total lineages used: `r nrow(barcodes)`

If there are any notes, they will appear here:

```{r}
#| label: data_notes
#| results: asis
#| fig-width: 8
#| fig-height: 10
if (!is.null(bioproj$data_notes)) {
    cat(bioproj$data_notes)
}

if (!is.null(bioproj$lit_notes)) {
    cat(bioproj$lit_notes)
}
```



## Provoc

```{r}
#| label: provoc
t0 <- Sys.time()
res <- provoc(count / coverage ~ .,
    data = coco,
    lineage_defs = barcodes,
    by = "sra"
)
```

ProVoC took `r round(difftime(Sys.time(), t0, units = "mins"), 2)` minutes.

```{r}
#| label: plot_provoc
res$date <- lubridate::ymd(res$date)

if (!"location" %in% colnames(res)) {
    res$location <- "only location"
}

gg <- autoplot(res, date_col = "date") + facet_wrap(~ location) +
    geom_smooth(formula = y ~ x, se = FALSE, method = "loess") +
    theme_bw()
suppressWarnings(print(gg))
```

